\textbf{Development of outdoor air pollution models for estimating exposure in the Barcelona Life Study Cohort (BiSC)}

### 2024-01-28 manuscript_BiSC_outdoor_AP.tex

@@

Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants is shown in \Cref{table1}. NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a median of  37.5 \(\pm\) 15.8 \( \mu \text{g/m}^3\). PM\textsubscript{2.5} observed levels ranged between 5.01 and 35.8 \( \mu \text{g/m}^3\) with a median of 11.3 \(\pm\) 7.14 \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign, ranging from 0.40 to 2.95 \(  \mu \text{g/m}^3\) with a median of 1.22 \(\pm\) 0.64 \( \mu \text{g/m}^3\). For PM\textsubscript{2.5} constituents  observed values during the BiSCAPE campaign varied substantially between elements with medians of 0.21 \(\pm\) 0.17 \(  \mu \text{g/m}^3\)  for  Fe,  6.36 \(\pm\) 4.61 \( \text{ng/m}^3 \) for Cu, and 28.2 \(\pm\) 27.8 \( \text{ng/m}^3 \) for Zn.

@@

\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using land use (LUR) regression models, dispersion models, and hybrid models. The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}. In general, LUR and HM models exhibit relatively good performance, while DM shown a more conservative performance evaluated by the $R^{2}$ and $RMSE$.

@@

%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure1.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\label{fig2}
\end{figure}

%%% --- Figure 4. Average exposure during the entire pregnancy period --- %%%
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9a\textwidth]{figures/combined_figure2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and hm(\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
\label{fig3}
\end{figure}

@@

%%% --- Table 2. Performance from external validation for LUR, DM, HM --- %%%
\begin{table}[h!]
\centering
    \caption{Performance from validation for LUR, DM, and HM models.}
    \label{table2}
    \begin{tabular}{llcccccc}
    \toprule
         Model & Validation & \multicolumn{6}{c}{R\textsuperscript{2 \textsubscript{ Validation}}} \\
         \cmidrule(lr){3-8}
         & & NO\textsubscript{2} & PM\textsubscript{2.5} & BC & PM\textsubscript{2.5 - Fe} & PM\textsubscript{2.5 - Cu} & PM\textsubscript{2.5 - Zn} \\
    \midrule
         LUR & External validation BiSC-Home & 0.56 & - & - & - & - & - \\
         DM & External validation BiSCAPE & 0.65 & 0.22 & 0.56 & - & - & - \\
         HM & External validation split (test data) & 0.64 & 0.65 & 0.69 & 0.65 & 0.56 & 0.36  \\
    \bottomrule
    \end{tabular}
\end{table}

Due to lack of data availability we only evaluate the performance of NO\textsubscript{2} for LUR models. The performance of the LUR model decreased in the external validation by 16\%. When validated DM against BiSCAPE measurements model performance increase for NO\textsubscript{2} but decrease for PM\textsubscript{2.5} and BC. HM showed consistent results in performance using both approaches to validate the models for NO\textsubscript{2} and PM\textsubscript{2.5} models showed a similar performance that the one obtained using the \textit{10-CV}. While for BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} performance varied between validation approaches by -19.7\%, +20.37\%, -20.0\% , and -18.8\% respectively.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three different approaches (LUR, ADM, and HM). Models show moderate to good performance 
evaluated by the defined metrics previously and displayed in more detail in the \textbf{supplementary materials}. As we expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed. We believe that by combining multiple data sources, and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.

>>>>

\subsection{Comparison between model estimates}
Exposure estimates obtained from LUR, DM, and HM were compared at the BiSC participant's address level for long- and short-term exposure. We compare the distribution of estimated concentrations between models and pollutants and calculate Spearman (Rho) correlation coefficients. We used scatterplots and Bland-Altmant plots to investigate the level of agreement between the different models.

@@

Distribution of the observed concentration for the BiSC-home and BiSCAPE campaigns for all pollutants are shown in \Cref{table1}. NO\textsubscript{2} concentration levels for the combination of BiSC-home and BiSCAPE campaigns ranged between 6.85 and 97.3 \(\mu \text{g/m}^3\) with a median of  37.5 \(\pm\) 15.8 \( \mu \text{g/m}^3\). PM\textsubscript{2.5} observed levels ranged between 5.01 and 35.8 \( \mu \text{g/m}^3\) with a median of 11.3 \(\pm\) 7.14 \(\mu \text{g/m}^3\), while BC levels showed less variation during the BiSCAPE campaign, ranging from 0.40 to 2.95 \(  \mu \text{g/m}^3\) with a median of 1.22 \(\pm\) 0.64 \( \mu \text{g/m}^3\). For PM\textsubscript{2.5} constituents  observed values during the BiSCAPE campaign varied substantially between elements with medians of 0.21 \(\pm\) 0.17 \(  \mu \text{g/m}^3\)  for  Fe,  6.36 \(\pm\) 4.61 \( \text{ng/m}^3 \) for Cu, and 28.2 \(\pm\) 27.8 \( \text{ng/m}^3 \) for Zn.

@@

\Cref{fig2} and \Cref{fig3} present maps of estimated exposures to NO\textsubscript{2}, PM\textsubscript{2.5}, BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} content  during the entire pregnancy for the BiSC participants using LUR, HM, and HM. The performance of the models varied substantially across different pollutants as shown in \Cref{fig2} and \Cref{fig3}. In general, LUR and HM models exhibit relatively good performance, while DM shows a more conservative performance evaluated by the $R^{2}$ and $RMSE$.

@@

%%% --- Figure 3. Average exposure during the entire pregnancy period --- %%%
% Put the figure text closer to the Figure 3
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure1_v2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period (\textit{NO$_2$}, \textit{PM$_{2.5}$}, \textit{BC}).}
\label{fig2}
\end{figure}

%%% --- Figure 4. Average exposure during the entire pregnancy period --- %%%
\captionsetup[figure]{skip=6pt}
% We add the figure with the study domain 
\begin{figure}[!htb]
\includegraphics[width=0.9\textwidth]{figures/combined_figure2_v2.png}
\caption{Outdoor air pollution exposure during the entire pregnancy period for LUR and HM (\textit{PM$_{2.5}$}, \textit{PM$_{2.5-Fe}$}, \textit{PM$_{2.5-Cu}$}, \textit{PM$_{2.5-Zn}$}.}
\label{fig3}
\end{figure}

@@

%%% --- Table 2. Performance from validation for LUR, DM, HM --- %%%
\begin{table}[h!]
\centering
    \caption{Performance from validation for LUR, DM, and HM models.}
    \label{table2}
    \begin{tabular}{llcccccc}
    \toprule
         Model & Validation & \multicolumn{6}{c}{R\textsuperscript{2 \textsubscript{ Validation}}} \\
         \cmidrule(lr){3-8}
         & & NO\textsubscript{2} & PM\textsubscript{2.5} & BC & PM\textsubscript{2.5 - Fe} & PM\textsubscript{2.5 - Cu} & PM\textsubscript{2.5 - Zn} \\
    \midrule
         LUR & External validation BiSC-Home & 0.56 & - & - & - & - & - \\
         DM & External validation BiSCAPE & 0.65 & 0.22 & 0.56 & - & - & - \\
         HM & External validation split (test data) & 0.64 & 0.65 & 0.69 & 0.65 & 0.56 & 0.36  \\
    \bottomrule
    \end{tabular}
\end{table}

Due to lack of data availability we only evaluate the performance of NO\textsubscript{2} for LUR models. The performance of the LUR model decreased in the external validation by 16\%. When validated DM against BiSCAPE measurements model performance increase for NO\textsubscript{2}, but decrease for PM\textsubscript{2.5} and BC. HM showed consistent results in performance using both approaches to validate the models for NO\textsubscript{2} and PM\textsubscript{2.5} models showed a similar performance that the one obtained using the \textit{10-CV}. While for BC, PM\textsubscript{2.5 - Fe}, PM\textsubscript{2.5 - Cu}, and PM\textsubscript{2.5 - Zn} performance varied between validation approaches by -19.7\%, +20.37\%, -20.0\% , and -18.8\% respectively.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% --- Discussion --- %%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
We successfully developed spatiotemporal models using three approaches (LUR, ADM, and HM). Models show moderate to good performance 
evaluated by previously defined metrics and displayed in more detail in the \textbf{supplementary materials}. As expected, hybrid models outperformed the LUR and DM models for the air pollutants assessed. We believe that by combining multiple data sources and integrating RF algorithm, we were able to capture non-linear and interactions between variables and consequently improve model predictive ability.
